### 2018-04-11 paper.Rmd
transcription -> transcript

abstract: |
  The Child Language Data Exchange System (CHILDES) has played a critical role in research on child language development. Access to these data can be complex for novices and difficult to automate for advanced users, however. To address these issues, we introduce childes-db, a database-formatted mirror of CHILDES that improves data accessibility and usability by offering novel interfaces, including browsable web applications and an R application programming interface (API). Along with a versioned infrastructure that facilitates reproducibility of past analyses, these interfaces lower barriers to analyzing naturalistic parent-child language, allowing for a wider range of researchers in language and cognitive development to easily leverage CHILDES in their work.

@@

Children's language acquisition is a core puzzle in cognitive science. How are the representations underlying the productive use of language learned over children's first years? Because a language is comprised of many individual pieces of information -- the names of particular objects, the marker for the past tense, the position of determiners -- language development is fundamentally a question of learning from environmental input, which is the only source of information on these specifics. Further, children are often reluctant experimental participants; often their most revealing and most sophisticated uses of language emerge in naturalistic play. Thus, language acquisition research must be grounded in the ecological behavior of children, both their language input and their language production in context.

The foundational resource for studies of language acquisition is recordings and transcripts of children. Starting with Roger Brown's [-@brown1973] work on Adam, Eve, and Sarah, longitudinal recordings of children's talk (and the talk they hear from their caregivers) have formed the basis for this field. But creating individual transcripts is extremely hard work. Beyond gathering the data, even orthographic transcriptions (without phonetic detail) can take ten times the length of the original recording to transcribe [@macwhinney2000]. Thus, researchers' desire for data from multiple ages or multiple children can very quickly outstrip their resources.

@@

Yet there are some challenges working with CHILDES, both for students and for advanced users. The CHILDES ecosystem uses a special file format (CHAT), which is stored as plain text but includes structured annotations grouped into tiers (stored on separate lines) so that information about utterances can be stored with accompanying information such as the phonological or morphological structure of the utterance. These files are analyzed using a command-line tool (CLAN) with a host of specific arguments that allow users to count word frequencies, compute statistics (e.g., mean length of utterance, or MLU). While this system is flexible and powerful, mastering the CHAT codes and the CLAN tool can be daunting for a beginner or student. These technical barriers decrease the ease of casual exploration by a novice researcher or in a classroom exercise.

On the opposite end of the spectrum, for data-oriented researchers who are interested in doing large-scale analyses of CHILDES, the current tools are also not ideal. CLAN software is an excellent tool for interactive exploration, but -- as a free-standing application -- it can be tricky to build into a processing pipeline written in a modern analytic language such as Python or R. Thus, researchers who would like to ingest the entire corpus (or some large subset) into a computational analyses typically write their own parsers of the CHAT format to extract the subset of the data they would like to use [e.g., @redington1998;@yang2013;@meylan2017].

The practice of writing custom parsers is problematic for a number of reasons. Effort is wasted in rewriting the same parser again and again. In addition, this process introduces a major source of errors and inconsistencies in data handling due to difficulties dealing with the many special cases in the CHAT standard. Further, these parsing scripts are rarely shared, leading to much greater difficulty in reproducing the exact numerical results from papers using CHILDES [see e.g., @meylan2017 for an example]. In addition, the CHILDES corpus itself is a moving target; computational work using the entire corpus at one time point may include a different set of data than subsequent work (due to new contributions or revisions) and there is no simple way for researchers to document exactly what version of the corpus has been used short of creating a full mirror of the data. These factors together lead to a lack of computational reproducibility, a major problem that keeps researchers from verifying or building on published research [@donoho2010;@stodden2016].

@@

We begin by describing the architecture of childes-db and the web applications that we provide. We next describe the `childesr` API, which allows for programmatic access to the data. We then present several worked examples of specific uses of the system for research and teaching using both the web app and the API.

@@

<!-- Remind readers why they're reading this: understanding the structure helps you use and understand the tools. SM: Moved this up from database format because it addresses more general issues -->
As described above, CHILDES is most often approached as a set of distinct CHAT files, which are then parsed by users (often using CLAN). 
In contrast to this parsing approach, which entails the sequential processing of strings, childes-db treats CHILDES as a set of linked tables, with records corresponding to intuitive abstractions such as words, utterances, and transcripts.
Users of data analysis languages like R or Julia, libraries like Pandas, or those familiar with Structured Query Language (SQL) will be familiar with operations on tables such as filtering (subsetting), sorting, aggregation (grouping), and joins (merges). 
These operations obviate the need for users to consider the specifics of the CHAT representation -- instead they simply request the entities they need for their research and allow the API to take care of the formatting details. 
We begin by orienting readers to the design of the system via a top-level description and motivation for the design of the database schema, details of the database’s current technical implementation, and our versioning scheme. 
Users primarily interested in accessing the database can skip these details and focus on access through the `childesr` API and the web apps. 
<!-- We first present the design and technical approach of childes-db, then describe the interactive analysis tools provided by our web-based frontend.  -->

@@

At its core, childes-db is a database consisting of a set of linked tabular data stores where records correspond to linguistic entities like words, utterances, and sampling units like transcriptions and corpora. 
<!-- what did you mean here? don't rows = records; SM: that's a description of a database in general; we can make a specific assertion about childes-db-->
The smallest unit of abstraction tracked by the database is a token, treated here as the standard (citation) orthographic form of a word.
Using the standardized written form of the word facilitates the computation of lexical frequency statistics, for comparison or aggregation across children or time periods. 
Deviations from the citation form -- reflecting the deviations from citation forms that are particularly common in the course of language development -- are kept as a separate (possibly null) field associated with each token.

Other tables in the database are hierarchical collections built out of tokens -- *utterance*, *transcription*, *corpus*, and *collection* -- which store attributes appropriate for each level of description.
Every entity includes attributes that link it to all higher-order collections, e.g., an utterance lists the transcript, corpus, and collection to which it belongs.
An *utterance* contains one or more words and includes fields such as the utterance type (declarative, question, etc.), total number of tokens, and the total number of morphemes (if the morphological structure is available in the original CHAT). 
A *transcription* consists of one or more utterances and includes the date collected, the name of the target child, and age in days (if defined), and the filename from CHILDES.
A *corpus* consists of one or more transcripts, corresponding to well-known collections like the Brown or Providence corpus.
Finally, a *collection* is a superordinate collection of corpora generally corresponding to a geographic region, following the convention in CHILDES.
Because every record can be linked to a top-level collection (generally corresponding to a language), each table includes data from all languages represented in CHILDES.

Participants (e.g., children) are represented separately from the token hierarchy because it is common for the same children to appear in multiple transcripts.
A participant identifier is associated with every word and utterance, including a name, role, 3-letter CHILDES identifier (CHI = child, MOT = mother, FAT = father, etc.), and the range of ages (or age of corresponding child) for which they are observed.
For non-child participants (caregivers and others), the record additionally contains an identifier for the corresponding target child to facilitate associating caregivers with children.

@@

childes-db is implemented as a MySQL database, an industry-standard, open-source relational database server that can be accessed directly from a wide range of programming languages.
The childes-db project provides hosted read-only databases for direct access and for `childesr` (described below) as well as compressed .sql exports for local installation.
While the former is appropriate for most users, local installation can provide performance gains (allowing a user to access the database on their local machine or on their network) as well as allowing users to store derived information in the same database.

In order to import the CHILDES corpora into the MySQL schema described above, it must first be accurately parsed and subsequently vetted to ensure its integrity.
We parse the XML (eXtensible Markup Language) release of CHILDES hosted by \url{childes.talkbank.org} using the NLTK library in Python [@bird2004].
<!-- details of our contribution to the CHILDES corpus module in NLTK---didn't we change it? -->
Logic implemented in Python converts the linear, multi-tier parse into a tabular format appropriate for childes-db.
This logic includes decisions regarding what information sources are captured in the current release of the database, and which are left for future development (detailed below).

The data imported into childes-db is subject to data integrity checks to ensure that our import of the corpora is accurate and preferable over ad-hoc parsers developed by many individual researchers.
<!-- MCF: more testing details would be useful -->
In order to evaluate our success in replicating CLAN parses, we compared unigram counts in our database with those outputted by CLAN, the command-line tool built specifically for analysis of transcripts coded in CHAT.
<!-- SM: what were the results of these comparisons?  -->
<!-- AS: technically this is all we do right now SM: what more can we do for low technical investment-->

@@

The content of CHILDES changes as additional corpora are added or transcriptions are updated; as of the time of writing, these changes are not systematically tracked.
We introduce a simple versioning system by adding a new complete parse of the current state of CHILDES every six months or as warranted by changes in CHILDES.
By default, users interact with the most recent version of the database available. 
But to support reproduction of results with previous versions of the database, we continue to host recent versions (up to the last three years / six versions) through our `childesr` API so that researchers can run analyses against specific historical versions of the database.
<!-- Potentially put some R code to show easy it is to access previous versions? MCF: Nah. -->
For versions more than three years old, we host compressed .sql files that users may download and serve using a local installation of MySQL server.
<!-- SM: How easy is it to implement versioning in the Shiny apps? This would show anyone using the web app the availability of the versions-->
<!-- SM: can people get to these sql files currently on the website?-->

@@

- Sparsely annotated tiers, e.g. phonology (`%pho`) and situation (`%sit`)
<!--- Media links, including timing information AS: we have media timestamps -->
<!--- Compound word markers, indicated with underscores ("green_eggs_and_ham") AS: we have this -->
- Tone direction and stress
- Filled pauses 
- Reformulations, word revision, and phrase revision, e.g. \textlt what did you\textgt [//] how can you see it ?
- paralinguistic material, e.g. [=! cries]

@@

We first discuss the childes-db web apps and then introduce the `childesr` `R` package.

## Interactive Web Apps

The ability to easily browse and explore the CHILDES corpora is a cornerstone of the childes-db project. 
To this end we have created powerful yet easy-to-use interactive web applications that enable users to visualize various dimensions of the CHILDES corpus: frequency counts, mean lengths of utterance, type-token ratios, and more. 
All of this is doable without the requirement of understanding command-line tools or any kind of programming knowledge as had been the case with CLAN.^[The LuCiD toolkit [@chang2017] provides related functionality for a number of common analyses. In contrast to those tools, which focus on filling gaps not covered by CLAN -- e.g., the use of n-gram models, incremental sentence generation, and distributional word classification -- our web apps focus on covering the same common tasks as CLAN, but making the outputs into browsable visualizations.]

Our web apps are built using Shiny, an R package that enables easy app construction using `R`. Underneath the hood, each web app is making calls to our `childesr` API and subsequently plotting the data using the popular R plotting package `ggplot2`. A user's only task is to configure exactly what should be plotted through a series of buttons, sliders, and text boxes. The user may specify what collection, corpus, child, age range, caregiver, etc., should be included in a given analysis. The plot is displayed and updated in real-time, and the underlying data are also available for download alongside the plot to facilitate customization. All of these analyses may also be reproduced using the `childesr` package.

The web apps are intended for the casual user who seeks to easy extract developmental indices quickly and without any technical overhead. We next walk through the three web apps provided on our website and describe their functionality in more detail.

@@

The lexical statistics of languaeg input to children have long been an object of study in child language acquisition research. Frequency counts of words in particular may provide insight into the cognitive, conceptual, and linguistic experience of a young child [see e.g., @ambridge2015 for review]. In this web app, inspired by ChildFreq [@baath2015], we provide users the ability to search for any word spoken by a participant in the CHILDES corpora and track the usage of that word by a child or caregiver over time. Because of the various toggles available to the user that can subset the data, a user may display frequency curves for a single child in the Brown corpus or all Spanish speaking children, if desired. In addition, users can plot frequency curves belonging to caregivers alongside their child for convenient side-by-side comparisons. A single word or multiple words may be entered into the input box.

@@

The syntactic complexity and lexial diversity of children's speech are similarly critical metrics for acquisition researchers [@miller1981,@watkins1995]. There are a number of well-established measures of children's speech that operationalize complexity and diversity. These developmental metrics have many applications in speech-language pathology (SLP), where measures outside of the normal range may be indicative of speech, language, or communication disorders.

Several of the most common of these measures are available in the Derived Measures app, which plots these measures across age for a given subset of data, again specified by collection, corpora, children, and speakers. As with the Frequency Counts app, caregivers' lexical diversity measures can be plotted alongside children's.

We have currently implemented:

- MLU-w (mean length of utterance in words, a common measure of complexity),^[Because of gaps in available CHILDES morphological data, MLU-w is computed and not MLU-m (mean length of utterance in morphemes). 
<!-- sm: I think we should go ahead and compute this for transcript with MOR tiers? Given that this was one of the first uses, I don't think Brian will like the 'underlying gaps' explanation. MCF: I agree. -->
In English, these two measures are highly correlated. <!-- sm: citation needed -->
alongside those of children.] 
- TTR (type-token ratio, a measure of lexical diversity), 
- MTLD (measure of textual lexical diversity), 
- HD-D (lexical diversity via the hypergeometric distribution), 
- and voc-d (a further measure of lexical diversity).  <!-- sm: MLU is not a measure of lexical diversity-->
<!-- MCF: need citations for voc-d HD-dand MTLD -->

\noindent As with the Frequency Counts app, a user may subset the data as he or she chooses, compare measures between caregivers and children, and aggregate across children from different corpora.

@@

Many times a researcher will want to investigate the statistics of a given corpus (i.e the size, number of utterances, number of tokens, etc.) before choosing a target corpus for their analysis. This web app is intended to provide a basic overview of various corpora in the CHILDES ecosystem, facilitating familiarity with the different corpora and thus assisting with the overall analytic end-to-end process.
<!-- note that the graphs don't currently include a grouping factor, so adding multiple corpora generates one line, so you can't actually compare two corpora in the same visualization -->
 <!-- cite original ChildFreq paper as inspiration? -->

<!-- figure of freq app -->

@@

Although the interactive analysis tools described above cover some of the most common use cases of CHILDES data, researchers interested in more detailed and flexible analyses will want to interface directly with the data in childes-db. Making use of the `R` programming language [@R-base], we provide the `childesr` package. `R` is an open-source, extensible statistical computing environment that is rapidly growing in popularity across fields and is increasing in use in child language research [e.g. @song2015; @norrman2015]. The `childesr` package abstracts away the details of connecting to and querying  the database. Users can take advantage of the tools developed in the popular `dplyr` package [@R-dplyr], which make manipulating large datasets quick and easy. We describe the commands that the package provides and then give several worked examples of using the package for analyses.

The `childesr` package is easily installed via CRAN, the comprehensive R archive network. To install, simply type: `install.packages("childesr")`. After installation, users have access to data loading functions that can be used to access different parts of the database:

@@

\noindent Each of these functions can be supplied with arguments that restrict the query to a particular subset of the data (e.g. by collection, by corpus, by speaker role, by target child age, etc.) and for a particular version of the data. For more detailed documentation, see the package repository (http://github.com/langcog/childesr).

>>>>

abstract: |
  The Child Language Data Exchange System (CHILDES) has played a critical role in research on child language development. However, current methods of using CHILDES data rely on a command line interface to extract developmental indices or custom parsers to extract data for larger-scale analyses. We introduce childes-db, a database-formatted mirror of CHILDES that improves data accessibility and usability by offering novel interfaces, including browsable web applications and an R application programming interface (API). Along with a versioned infrastructure that facilitates reproducibility of past analyses, these interfaces lower barriers to analyzing naturalistic parent-child language, allowing for a wider range of researchers in language and cognitive development to easily leverage CHILDES in their work.

@@

Children's language acquisition is a core puzzle in cognitive science. How are the representations underlying the generative use of language learned over children's first years? Because a language is comprised of many individual pieces of information -- the names of particular objects, the marker for the past tense, the position of determiners -- language development is fundamentally a question of learning from environmental input, which is the only source of information on these specifics. Further, children are often reluctant experimental participants; often their most revealing and most sophisticated uses of language emerge in naturalistic play. Thus, language acquisition research must be grounded in the ecological behavior of children, both their language input and their language production in context.

The foundational resource for studies of language acquisition is recordings and transcripts of children. Starting with Roger Brown's [-@brown1973] work on Adam, Eve, and Sarah, longitudinal recordings of children's talk (and the talk they hear from their caregivers) have formed the basis for this field. But creating individual transcripts is extremely hard work. Beyond gathering the data, even orthographic transcriptions (without phonetic detail) can take ten times the length of the original recording to transcribe. Thus, researchers' desire for data from multiple ages or multiple children can very quickly outstrip their resources.

@@

Yet there are some challenges working with CHILDES, both for students and for advanced users. The CHILDES ecosystem uses a special file format (CHAT), which is stored in a plain text format but includes structured annotations grouped into tiers (stored on separate lines) so that information about utterances can be stored with accompanying information such as the phonological or morphological structure of the utterance. These files are analyzed using a command-line tool (CLAN) with a host of specific arguments that allow users to count word frequencies, compute statistics (e.g., mean length of utterance, or MLU). While this system is flexible and powerful, mastering the CHAT codes and the CLAN tool can be daunting to navigate for a beginner or student. These technical barriers decrease the ease of casual exploration by a novice researcher or in a classroom exercise.

On the opposite end of the spectrum, for data-oriented researchers who are interested in doing large-scale analyses of CHILDES, the current tools are also not ideal. CLAN software is an excellent tool for interactive exploration, but -- as a free-standing application -- can be tricky to build into a processing pipeline written in a modern analytic platform such as Python or R. Thus, researchers who would like to ingest the entire corpus (or some large subset) into a computational analysis typically write their own parsers of the CHAT format to extract the subset of the data they would like to use [e.g., @redington1998;@yang2013;@meylan2017].

This practice is problematic for a number of reasons. Effort is wasted in rewriting the same parser again and again. In addition, this process introduces a major source of errors and inconsistencies in data handling due to difficulties dealing with the many special cases in the CHAT standard. Further, these scripts are rarely shared, leading to much greater difficulty in reproducing the exact numerical results from papers using CHILDES [see e.g., @meylan2017]. This lack of computational reproducibility is widely recognized as problematic because it does not allow researchers to verify or build on published research [@donoho2010;@stodden2016]. Finally, the CHILDES corpus itself is a moving target;  computational work using the entire corpus at one time point may include a different set of data than subsequent work (due to new contributions or revisions).

@@

We begin by describing the architecture of childes-db and the web applications that we provide. We next describe the `childesr` API, which allows for programmatic access to the data. We then present two worked examples of specific uses of the system using both the web app and the API. The first examines the frequency of color words in children's input; the second quantifies the growth of lexical diversity across genders. We end by presenting some suggestions for the use of the childes-db web front-end in classroom teaching.

@@

<!-- Remind readers why they're reading this: understanding the structure helps you use and understand the tools. SM: Moved this up from database format because it addresses more general issues -->
Unlike the *parsing* approach in CLAN which entails the sequential processing of strings, childes-db treats CHILDES as a set of linked tables, with records corresponding to intuitive abstractions such as words, utterances, and transcripts.
Users of data analysis languages like R or Julia, libraries like Pandas, or those familiar with Structured Query Language (SQL) will be familiar with operations on tables such as filtering (subsetting), sorting, aggregation (grouping), and joins (merges).
Although we expect most users to interact with the database through the `childesr` API for the `R` language described below, we begin with a top-level description and motivation for the design of the database schema, details of the database’s current technical implementation, and our versioning scheme.
<!-- We first present the design and technical approach of childes-db, then describe the interactive analysis tools provided by our web-based frontend.  -->

@@

At its core, childes-db is a database consisting of a set of linked tabular data stores where rows correspond to linguistic entities like words, utterances, and sampling units like transcripts and corpora. 
<!-- what did you mean here? don't rows = records; SM: that's a description of a database in general; we can make a specific assertion about childes-db-->
The smallest unit of abstraction tracked by the database is a token, treated here as the citation orthographic form of a word.
Using the standardized written form of the word facilitates the computation of lexical frequency statistics, for comparison or aggregation across children or time periods. 
Deviations from the citation form -- reflecting the deviations from citation forms that are particularly common in the course of language development -- are kept as a separate (possibly null) field associated with each token.

Other tables in the database comprise a hierarchical system of collections built on top of tokens -- *utterance*, *transcript*, *corpus*, and *collection* -- which stores attributes appropriate for each level of description.
Every entity includes attributes that link it to all higher-order collections, e.g., an utterance lists the transcript, corpus, and collection to which it belongs.
An *utterance* contains one or more words and includes fields such as the utterance type (declarative, question, etc.), total number of tokens, and the total number of morphemes (if the morphological structure is available in the original CHAT).

A *transcript* consists of one or more utterances and includes the date collected, the name of the target child, and age in days (if defined), and the filename from CHILDES.
A *corpus* consists of one or more transcripts, corresponding to well-known collections like the Brown or Providence corpus.
Finally, a *collection* is a superordinate collection of corpora generally corresponding to a geographic region, following the convention in CHILDES.
Because every record can be linked to a top-level collection (generally corresponding to a language), each table includes data from all languages represented in CHILDES.

We chose to treat participants separate from the above hierarchy because children can occur across transcripts (e.g., the Brown Corpus).
A participant identifier is associated with every word and utterance, including a name, role, 3-letter CHILDES identifier (CHI, MOT, FAT, etc.), and the range of ages (or age of corresponding child) for which they are observed.
For non-child participants (caregivers and others), the record additionally contains an identifier for the corresponding target child to associate caregivers and children.

@@

The above tabular datastore is implemented as a MySQL database.
As an industry-standard, open-source relational database server, the above database can be accessed directly from a wide range of programming languages.
The childes-db project provides hosted read-only databases for direct access and for `childesr` (described below) as well as compressed .sql exports for local installation.
While the former is appropriate for most users, local installation can provide performance gains (allowing a user to access the database on their local machine or on their network) as well as allowing users to store derived information in the same database.

In order to import the CHILDES corpora into the MySQL schema described above, it must first be accurately parsed and subsequently vetted to ensure its integrity.
We parse the XML (eXtensible Markup Language) release of CHILDES hosted by \url{childes.talkbank.org} using the NLTK library in Python.
<!-- citation for NLTK -->
<!-- details of our contribution to the CHILDES corpus module in NLTK---didn't we change it? -->
Logic implemented in Python converts the linear, multi-tier parse into a tabular format appropriate for childes-db.
This logic includes decisions regarding what information sources are captured in the current release of the database, and which were left for future development (detailed below).

<!-- testing details -->
The data imported into childes-db requires extensive data integrity checks to ensure that our parse of the corpora is accurate and preferable over ad hoc parsers developed by many individual researchers.
In order to evaluate this, we compared unigram counts in our database with those outputted by CLAN, the command-line tool built specifically for analysis of transcripts coded in CHAT.
<!-- SM: what were the results of these comparisons?  -->
<!-- AS: technically this is all we do right now SM: what more can we do for low technical investment-->

@@

The content of CHILDES changes as additional corpora are added or transcriptions are updated; as of the time of writing, these changes are not systematically tracked.
Here we introduce a simple versioning system by adding a new complete parse of the current state of CHILDES every six months.
By default, users interact with the most recent version of the database available. 
To support reproduction of results with previous versions of the database, we continue to host recent versions (up to the last three years / six versions) through our `childesr` API so that researchers can run analyses against specific historical versions of the database.
<!-- Potentially put some R code to show easy it is to access previous versions? -->
For versions more than three years old, we host compressed .sql files that users may download and serve using a local installation of MySQL server.
<!-- SM: How easy is it to implement versioning in the Shiny apps? This would show anyone using the web app the availability of the versions-->
<!-- SM: can people get to these sql files currently on the website?-->

@@

- Sparsely annotated tiers, e.g. phonology (`%pho`) and situation (`%sit`)
<!--- Media links, including timing information AS: we have media timestamps -->
<!--- Compound word markers, indicated with underscores ("green_eggs_and_ham") AS: we have this -->
- Tone direction and stress
- Filled pauses 
- Reformulations, word revision, and phrase revision, e.g. <what did you> [//] how can you see it ?
- paralinguistic material, e.g. [=! cries]

@@

We first discuss the childes-db web apps and then introduce the `childesr` package.

## Interactive Analysis Tools

<!-- need a good shorthand for the interactive analysis tools: web apps -->

The ability to easily browse and explore the CHILDES corpora is a cornerstone of the childes-db project. To this end we have created powerful yet easy-to-use interactive web applications that enable users to visualize various dimensions of the CHILDES corpus: frequency counts, mean lengths of utterance, type-token ratios, and more. All of this is doable without the requirement of understanding command-line tools or any kind of programming knowledge as had been the case with CLAN. The only true requirements are a browser and internet connection.

The interactive plots (web apps) are built using Shiny, an R package that enables easy building of web applications in R. Underneath the hood, each web app is making calls to our `childesr` API and subsequently plotting the data using the R plotting package `ggplot2`. A user's only task is to configure exactly what should be plotted through a series of buttons, sliders, and text boxes. The user may specify what collection, corpus, child, age range, caregiver, etc., should be included in a given analysis (e.g. mean length of utterance). The plot is displayed and updated in real-time, and if a user were to feel the need to explore the data independently then that data is also available for download alongside the plot. All of these analyses may also be reproduced only using the `childesr` package; the web apps were intended for the casual user who seeks to easy extract developmental indices without any technical overhead.

In the following sections we will walk through the three web apps made available on our website and describe their functionality in more detail.

@@

A child's lexical statistics has long been an object of study in child language acquisition research. Namely, the frequency of a word <!-- something important about frequency and development?-->. In this web app, we provide users the ability to search for any word or string

The study of frequency counts of individual lexical items as long been the

<!-- sm: intro sentence here for the role of frequency in CLA research  --> 
The Frequency Counts app allows users to enter a word (or words) in a text box and see the frequency (in parts per million) in which this word was uttered across a child's age for a particular subset of the data, whether it is an individual child in the Brown corpus or all Spanish speaking children and their mothers. <!-- sm: "whether..." is awk.  -->
The ability to view the caregiver’s usage of a term alongside a child’s assists researchers in better understanding the relationship between input and uptake. Users can also plot word trajectories alongside each other (such as “mom” and “dad” or “front” and “back”) for convenient cross-word comparisons.

@@

<!-- sm: intro sentence here for the role of developmental metrics in CLA research and in downstream applications like SLP--> 
There are a number of ways to measure the lexical diversity of an individual's productive speech. <!-- sm: different intro; MLU isn't a lexical diversity measure -->
Some of these measures, including MLU-w (mean length of utterance in words), TTR (type-token ratio), MTLD (measure of textual lexical diversity), and voc-d are available in the Derived Measures app, which plots these measures across age for a given subset of data, again specified by collection, corpora, children, and speakers. <!-- sm: MLU is not a measure of lexical diversity-->
Because of the underlying gaps in available CHILDES morphological data, MLU-w is computed and not MLU-m (mean length of utterance in morphemes). <!-- sm: I think we should go ahead and compute this for transcript with MOR tiers? Given that this was one of the first uses, I don't think Brian will like the 'underlying gaps' explanation-->
In the English language, these two measures have been shown to be highly correlated. <!-- sm: citation needed -->
As with the Frequency Counts app, caregiver’s lexical diversity measures can be plotted alongside those of children.

@@

The population viewer is a tool that helps users understand the relative sizes and temporal extents of various corpora by displaying the total number of utterances or tokens in the available corpora at particular ages. In many instances the size of a given corpus may inform researchers’ decisions as to what is the best target dataset for analysis. <!-- note that the graphs don't currently include a grouping factor, so adding multiple corpora generates one line, so you can't actually compare two corpora in the same visualization -->

@@

Although the interactive analysis tools described above cover some of the most common use cases of CHILDES data, researchers interested in more detailed and flexible analyses will want to interface directly with the data in childes-db. Making use of the `R` programming language [@R-base], we provide the `childesr` package to help researchers accomplish this task. `R` is an open-source, extensible statistical computing environment that is rapidly growing in popularity across fields and is increasing in use in child language research [e.g. @song2015; @norrman2015]. The `childesr` package abstracts away the details of connecting to <!-- sm: and querying-->  the database. Users can take advantage of the tools developed in the popular `dplyr` package [@R-dplyr], which make manipulating large datasets quick and easy. We describe the commands that the package provides and then give several worked examples of using the package for analyses.

The `childesr` package is easily installed via CRAN, the comprehensive R archive network. To install, simply type: `install.packages("childesr")`. After installation, users have access to data loading functions that retrieve a variety of subsets <!-- sm: entities? variety of subsets sounds non-hierarchical-->  of the data:

@@

Each of these functions can be supplied with arguments that restrict the query to a particular subset of the data (e.g. by collection, by corpus, by speaker role, by target child age, etc.) For more detailed documentation, see the package repository (http://github.com/langcog/childesr).
